\mytitle

### 2023-05-10 chapters/depsynth/impl.tex
abstract; more depsynth

\paragraph{Decomposing crash consistency queries.}
As \cref{todo} explains, \depsynth offloads a query to an SMT
solver when it first encounters a new litmus test to determine
whether or not the current set of dependency rules satisfy the
crash consistency property for that test.
In practice, this query is encoded by representing the \textit{committed status}
of each write from a litmus test as a boolean variable
and each edge in the happens-before graph as an implication between those variables.
% The value of these variables also affects the disk state,
% which is used
For tests that generate a large happens-before graph
with respect to the current dependency rules,
the query can take tens of seconds.

Experimentally, we observed that for these queries with long runtimes,
decomposing the query by concretizing one write-committed variable
drastically speeds up overall solving time
(at most \splitvariablebestspeedup \times speedup for a single litmus test).
Specifically, considering the set of write-committed variables $V$,
the happens-before graph $G\subset V\times V$,
and the original encoding for the litmus test crash consistency check $P(G)$,
we choose a variable $v_0\in V$ and issue the following two queries:
\begin{enumerate}
  \item $v_0 \wedge \forall v\in V. (v, v_0)\in G \rightarrow v) \wedge P(G)$
  \item $\neg v_0 \wedge \forall v\in V. (v, v_0)\in G \rightarrow \neg v) \wedge P(G)$
\end{enumerate}
The litmus test is crash consistent with respect to the dependency rules
if and only if both queries succeed.

Intuitively, this decomposition can result in faster solving times for two reasons.
First, this optimization can be seen as a pre-processing variable selection step.
By taking advantage of the relationship between nodes in the happens-before graph
(which is abstracted away in the constraints sent to the solver),
we are effectively telling the solver to select a good first variable for branching.
Second, concretizing write-committed variables can also indirectly simplify the
crash consistency predicate for the litmus test.
The crash consistency predicate is a function over disk states,
and the disk state itself depends on the write-committed variables.
Thus, by assigning concrete values to some of those variables,
constraints representing the crash consistency predicate over the disk state
may be simplified or eliminated.

\paragraph{Concretizing More Variables.}
%Specifically, we choose a variable whose corresponding 
Upon further experimentation, we noticed both that the number of
concretized variables and the specific variables chosen for concretization
significantly impacted solving time.
Ultimately, % TODO chose log n vars such that the most edges were eliminated

>>>>

\paragraph{Decomposing crash consistency queries.}
As \cref{sec:alg:alg} explains, \depsynth offloads a query to an SMT
solver when it first encounters a new litmus test to determine
whether or not the current set of dependency rules satisfy the
crash consistency property for that test.
In practice, this query is encoded by representing the \textit{committed status}
of each write from a litmus test as a boolean variable
and each edge in the happens-before graph as an implication between those variables.
% The value of these variables also affects the disk state,
% which is used
For tests that generate a large happens-before graph
with respect to the current dependency rules,
the query can take tens of seconds.

Experimentally, we observed that for these queries with long runtimes,
decomposing the query by concretizing one or more write-committed variables
drastically speeds up solving time.
(averaging $\splitvariableavgspeedup \times$ and at most $\splitvariablebestspeedup \times$ speedup for a single litmus test).
Specifically, considering the set of write-committed variables $V$,
the happens-before graph $G\subset V\times V$,
and the original encoding for the litmus test crash consistency check $P(G)$,
we choose a variable $v_0\in V$ and issue the following two queries:
\begin{enumerate}
  \item $v_0 \wedge \forall v\in V. (v, v_0)\in G \rightarrow v) \wedge P(G)$
  \item $\neg v_0 \wedge \forall v\in V. (v, v_0)\in G \rightarrow \neg v) \wedge P(G)$
\end{enumerate}
The litmus test is crash consistent with respect to the dependency rules
if and only if both queries succeed.
This concretization can be repeated for any number of variables,
resulting in $2^n$ queries for $n$ concretized variables.
Ultimately, this optimization gives a speedup of $\splitvariableoverallspeedup \times$
for the \shardstore case study discussed in \cref{sec:eval:shardstore},
reducing end-to-end runtime from \shardstoresynthesistime minutes to \splitvariablenewtime minutes.

\paragraph{Choosing Variables to Concretize.}
Intuitively, this decomposition results in faster solving times for two reasons.
First, this optimization can be seen as a pre-processing variable selection step.
By taking advantage of the relationship between nodes in the happens-before graph
(which is abstracted away in the constraints sent to the solver),
we are effectively telling the solver to select a good first variable for branching.
Second, concretizing write-committed variables can also indirectly simplify the
crash consistency predicate for the litmus test.
The crash consistency predicate is a function over disk states,
and the disk state itself depends on the write-committed variables.
Thus, by assigning concrete values to some of those variables,
constraints representing the crash consistency predicate over the disk state
may be simplified or eliminated.

With this intuition in mind, we chose concretization variables
that will most effectively simplify the constraints given to the solver.
We found that for a litmus test with $N$ writes,
choosing $log(N)$ variables to concretize gave the best results.
Moreover, the choice of which variables to concretize is determined by the
happens-before graph over the litmus test writes
(w.r.t. \depsynth's current candidate rules set).
Specifically, we choose the $log(N)$ variables whose corresponding
node in the happens-before graph has the most ancestors and descendants.
% TODO elaborate on this?

### 2023-05-10 chapters/depsynth/results.tex
abstract; more depsynth

\subsection{Other Storage Systems}\label{sec:eval:other}

In addition to the \shardstore case study, % in \cref{sec:eval:shardstore},
we have also used \depsynth
to implement a log-structured file system~\cite{rosenblum:lfs}.
%with support for files and directories.
The file system supports five standard POSIX operations:
\texttt{open}, \texttt{creat}, \texttt{write}, \texttt{close}, and \texttt{mkdir}.
While our implementation is simple (300 lines of Racket code) compared to production file systems,
it has metadata structures for files and directories,
and so has its own subtle crash consistency requirements.
For example, updates to data and inode blocks must reach the
disk before the pointer to the tail of the log is updated.
To synthesize dependency rules for this file system,
we randomly generated \lfsinputtests{} litmus tests
with at most \lfsmaxops{} operations.
\depsynth synthesized a set of \lfsnumrules{} dependency rules
in \lfssynthesistime{} to make the file system crash consistent,
and during the search,
invokes \textsc{RulesForTest} for only \lfstestsused{} tests.
This result shows that \depsynth can automate crash consistency for storage systems other than key-value stores.

>>>>

\subsection{Other Case Studies}\label{sec:eval:other}
% TODO include extra paragraph here, from the ECOOP edits

In addition to the \shardstore case study, % in \cref{sec:eval:shardstore},
we have applied \depsynth to two other storage systems.
The first is a modification of \shardstore with a different underlying disk hardware model.
The second is a simple log-strucutred file system.
\depsynth was effectively able to synthesize rules for both of these
additional cases, demonstrating the generality of our approach.

\paragraph{\shardstore with SMR.}
Part of \depsynth's requirement for input storage system implementations 
is a model of how the underlying disk hardware behaves.
For example, some storage devices automatically enforce
an order over writes or may disallow particular orders of writes
to the same extent on disk.
In the case study of \shardstore discussed above,
we use a disk model that represents a conventional magnetic-recording (CMR) hard disk.
For this second application of \depsynth, we instead use a model of
shingled magnetic recording (SMR) disks and synthesize rules
for \shardstore under this new model.

To understand how this difference in disk models will affect the synthesized rules,
we first give a brief overview of SMR disks.
Like traditional CMR disks, SMR drives store data on platters that contain tracks
read by a magnetic needle.
Where the technologies differ is in the layout of the tracks:
tracks in CMR disks are non-overlapping and lay parallel to each other,
while tracks in SMR lay partially over one another.
This allows SMR disks to store data more densely than their CMR counterparts,
but requires that all writes to a single SMR extent be \textit{appends}.
In order to write data at a location before the append pointer in an SMR extend,
the whole extent must first be wiped.

%This has two implications for our use of \depsynth over \shardstore with
%an SMR disk model.
%First, since SMR drives will force data to be appended to a single extent \textit{in order},
%rules that specify an order over writes to the same extent that \textit{matches the append order} are not necessary.
%Second, since data cannot be written out-of-order to an SMR extent without first wiping the extent,
%rules that specify any order over writes to the same extent \textit{other than the append order}
%will not be enforceable.
% TODO conclude this?

We ran \depsynth to synthesize rules for the \shardstore-like system with an underlying SMR disk model.
To do so, we generated a new set of \shardstoreinputtests litmus tests, each with a maximum
of \shardstoremaxops operations.
For this new system, \depsynth generated \shardstoresmrnumrules
in \shardstoresmrsynthesistime minutes.
During this process, \depsynth invoked the \rulesfortest procedure
for \shardstoresmrtestsused of the total \shardstoreinputtests tests.
%Note that though fewer rules were synthesized in a short amount of time
%compared to the case study in \cref{sec:eval:shardstore},
%this second synthesis task was not simply a sub-task of the first --- 
These results demonstrate that \depsynth is resiliant to changes in hardware used by storage systems.

\paragraph{Log-Structured File System.}
We have used \depsynth
to implement a log-structured file system~\cite{rosenblum:lfs}.
%with support for files and directories.
The file system supports five standard POSIX operations:
\texttt{open}, \texttt{creat}, \texttt{write}, \texttt{close}, and \texttt{mkdir}.
While our implementation is simple (300 lines of Racket code) compared to production file systems,
it has metadata structures for files and directories,
and so has its own subtle crash consistency requirements.
For example, updates to data and inode blocks must reach the
disk before the pointer to the tail of the log is updated.
To synthesize dependency rules for this file system,
we randomly generated \lfsinputtests{} litmus tests
with at most \lfsmaxops{} operations.
\depsynth synthesized a set of \lfsnumrules{} dependency rules
in \lfssynthesistime{} to make the file system crash consistent,
and during the search,
invokes \textsc{RulesForTest} for only \lfstestsused{} tests.
This result shows that \depsynth can automate crash consistency for storage systems other than key-value stores.

### 2023-05-10 chapters/intro.tex
abstract; more depsynth

% What is hard?
In recent years, the development of software systems has become increasingly
complex.
From the construction of large production key-value stores to Linux kernel JIT compilers
% TODO should I include these examples?
to the design of blockchain protocols and fault-tolerent networks,
modern systems must satisfy ever-increasing requirements on performance.
Additonally, the
correctness of these systems is now more important than ever. The
afformentioned storage systems are responsible for safely storing the data
of millions of people, while
the incorrect behavior of in-kernel compilers can subject the tens of millions of
Linux users to a slew of vulnerabilities.

@@

\cref{c:depsynth} presents \depsynth, a tool for automating
crash consistency in storage systems. In \cfref{c:depsynth},
we introduce a new dependency rule language that describes rules
which specify required write-before relationships for a storage system
at runtime. This abstraction over write-before relationships, along
with a new per-litmus-test search, allows \depsynth to synthesize
correct dependency rules for a production system.
Moreover, these generated rules have nearly the same effect on performance
as dependencies hand-written by experts.

>>>>

% What is hard?
In recent years, the development of software systems has become increasingly
complex.
From the construction of large production key-value stores to Linux kernel JIT compilers,
% TODO should I include these examples?
%to the design of blockchain protocols and fault-tolerent networks,
modern systems must satisfy ever-increasing requirements on performance.
Additonally, the
correctness of these systems is now more important than ever. The
afformentioned storage systems are responsible for safely storing the data
of millions of people, while
the incorrect behavior of in-kernel compilers can subject the tens of millions of
Linux users to a slew of vulnerabilities.

@@

\cref{c:depsynth} presents \depsynth, a tool for automating
crash consistency in storage systems. In \cref{c:depsynth},
we introduce a new dependency rule language that describes rules
which specify required write-before relationships for a storage system
at runtime. This abstraction over write-before relationships, along
with a new per-litmus-test search, allows \depsynth to synthesize
correct dependency rules for a production system.
Moreover, these generated rules have nearly the same effect on performance
as dependencies hand-written by experts.
